Notes on Control Plane Inference

### 2021-04-15 inference.tex
simplify

% record fields
\newcommand{\matches}{\mathsf{matches}}
\newcommand{\action}{\mathsf{action}}
\newcommand{\actions}{\mathsf{actions}}
\newcommand{\keys}{\mathsf{keys}}
\newcommand{\data}{\mathsf{data}}
\newcommand{\params}{\mathsf{params}}
\newcommand{\id}{\mathsf{id}}

% commands
\newcommand{\assert}{\mathop{\mathsf{assert}}}
\newcommand{\assume}{\mathop{\mathsf{assume}}}
\newcommand{\apply}{\mathsf{apply}}
\newcommand{\choiceop}{\rotatebox[origin=c]{90}{$\sqsubset\!\!\!\sqsupset$}}
\newcommand{\choice}{\mathbin{\choiceop}}
\DeclareMathOperator*{\bigchoice}{\scalerel*{\choiceop}{\sum}}
\newcommand{\havoc}{\texttt{<?>}}
\newcommand{\SKIP}{\mathsf{skip}}

% math
\newcommand{\argmin}{\mathop{\textrm{argmin}}}
\newcommand{\WP}{\textit{wp}}
\newcommand{\satisfy}[3]{({#1,#3}) \models_{#2}}
\newcommand{\notsatisfy}[3]{({#1,#3})\not\models_{#2}}
\newcommand{\nondet}{\mathsf{nondet}}
\newcommand{\ghostly}{\mathsf{ghostly}}
\newcommand{\state}[1]{\llparenthesis#1\rrparenthesis}

@@

\begin{figure}[htp]
  \[\begin{array}{lclll}
    \multicolumn{3}{l}{\Value} \\
    v & ::= & [n]_w  & \textit{Bitvector Literal} & (a,g,p,\iota) \\
    \multicolumn{3}{l}{\SymbRow} \\
    \rho & ::= & \left\{
    \begin{array}{l}
      \matches = \List[x];\\
      \action = x;\\
      \data = \List[x]
    \end{array}\right\} & \textit{Symbolic Row} & (\iota)\\
    \multicolumn{3}{l}{\BVExpr} \\
    e & ::= & v & \textit{Value} & (a,g,p,\iota)\\
      & \mid & h.f & \textit{Header Access} & (a,g,p,\iota)\\
      & \mid & \havoc & \textit{Havoc} & (a,g,p,\iota) \\
      & \mid & \rho.\matches[i] & \textit{Symbolic Match} & (\iota)\\
      & \mid & \rho.\action & \textit{Symbolic Action} & (\iota) \\
      & \mid & \rho.\data[i] & \textit{Symbolic Action Data} & (\iota) \\
      & \mid & x & \textit{Action Data Variable} & (a^*) \\
      & \mid & e \binop e & \textit{Arbitrary Binary Operation} & (a,g,p,\iota) \\
    & \mid & \unop e & \textit{Arbitrary Unary Operation} & (a,g,p,\iota) \\
    \multicolumn{3}{l}{\Table} \\
    t & ::= & \left\{\begin{array}{l}
    t.\id = v; \\
    t.\keys = \List[h.f]; \\
    t.\actions = \List[\lambda \vec x.\,a]\\
    \end{array}\right\} & \textit{Table} & (p)\\
    \multicolumn{3}{l}{\BExpr} \\
    b & ::= & \FALSE  & \textit{Absurdity}& (a,g,p,\iota)\\
      & \mid & b \Rightarrow b & \textit{Implication} & (a,g,p,\iota)\\
      & \mid & e = e & \textit{Bitvector Equality} & (a,g,p,\iota)\\
    & \mid & \forall \rho \in t.~b & \textit{Row Universality} & (\iota)\\
    \multicolumn{4}{l}{c \in \Cmd, a \in \Action, g \in \GCL, p \in \Prog, \iota \in \Instr } \\
    c & ::=  & h.f := e & \textit{Assignment} & (g,p,\iota) \\
      & \mid & \assert b & \textit{Assertion} & (g,p,\iota)\\
      & \mid & \assume b & \textit{Assumption} & (g,p,\iota) \\
      & \mid & t.\apply() & \textit{Table Application} & (p)\\
      & \mid & c;c & \textit{Sequential Composition} & (a,g,p,\iota) \\
      & \mid & c \choice c & \textit{Nondeterministic Choice} & (g,p,\iota)\\
      & \mid & \{c\} & \textit{Scope} & (g,p,\iota)\\
    \multicolumn{3}{l}{
      \begin{array}{ll}
        x \in \Var & h \in \Hdr\\
        n,w,i \in \mathbb{N} & f \in \Field
    \end{array}} & \textit{Sets }\\
  \end{array}
\]
\caption{Grammar of Commands $c \in \Cmd$. Legal action indicated by $a \in
  \Action$, legal instrumented command variants indicated by $\iota \in \Instr$,
  dataplane programs indicated by $p \in \Prog$, and table-free commands
  indicated by $g \in \GCL$.}
\label{fig:grammar}
\end{figure}

@@

Second, $p \in \Prog \subset \Cmd$ represents the expected set of programs extracted from a
P4 interface, most notably, it contains tables $t$ and no symbolic selection
statements or symbolic expersions.

Conversely $\iota \in \Instr \subset \Cmd$ represents a program instrumented
with symbolic variables. A program $\iota$ cannot apply tables using
$t.\apply()$, but instead uses havoc statements, to
introduce symbolic variables $\rho$ in $c$.

Finally $g \in \GCL \subset \Prog \subset \Cmd$ represents the core algebraic
flavor of GCL without havoc, symbolic tables, or table application.

@@

A state map $\sigma : \Value \to \Row^*$ is well-formed for a command $c$ iff the
following conditions hold:
\begin{enumerate}
\item[\textsc{WellFormedRows}] For every $t\in \tables(c)$, every row in $\sigma(t.\id)$ is well-formed.
\item[\textsc{NoShadow}] For every $t \in \tables(c)$, with $r_1,\ldots,r_n =
  \sigma(t.\id)$, for every $1 \leq i \leq n$, there does not exist $1 \leq j <
  i$ such that $r_i.\matches = r_j.\matches$.
\end{enumerate}
\paragraph{Row Models}
We also need models $\mu \in \Model$ to indicate which symbolic rows corresponds
to which concrete rows, i.e. $\Model = \SymbRow \rightharpoonup \Row$. We write
$\mu\{\rho \mapsto r\}$ to extend $\mu$ by mapping $\rho$ to $r$, and write $\mu
\setminus \rho$, to remove $\rho$ from the map.

@@

\begin{figure}[htp]
  \[\begin{array}{l >{\triangleq}cl}
  \multicolumn{3}{l}{\denote{c}^\sigma : \Model \to \Pkt \to \mathcal{P}(\Model \times \Pkt)_\error }\\
    \denote{h.f := e}^\sigma\,\mu\,\pkt && \{(\mu, \pkt\{h \mapsto v\} \mid v \in \edenote{e}\,\mu\,\pkt\})\} \\
    \denote{\assume b}^\sigma\,\mu\,\pkt &&
    \begin{cases}
      \{(\mu,\pkt)\} & \satisfy \sigma \pkt \mu b \\
      \{\} & \textit{otherwise} \\
    \end{cases}\\
    \denote{\assert b}^\sigma\,\mu\,\pkt &&
    \begin{cases}
      \{(\mu,\pkt)\} & \satisfy \sigma \pkt \mu b \\
      \error & \textit{otherwise}
    \end{cases} \\
    \denote{c_1;c_2}^\sigma\,\mu\,\pkt&&
    \bigcup_{(\mu',\pkt') \in \denote{c_1}^\sigma\,\mu\,\pkt}\denote{c_2}^\sigma\,\mu'\,\pkt' \\
    \denote{c_1 \choice c_2}^\sigma\,\mu\,\pkt &&
    \denote{c_1}^\sigma\,\mu\,\pkt \cup \denote{c_2}^\sigma\,\mu\,\pkt \\
    \denote{\{c\}}^\sigma\,\mu\,\pkt &&
    \{(\mu,\pkt') \mid (\mu',\pkt') \in \denote{c}^\sigma\,\mu\,\pkt \}\\
    \denote{t.\apply()}^\sigma\,\mu\,\pkt &&
    \begin{cases}
      \denote{c[\vec d/\vec x]}^\sigma\,\mu\,\pkt,
      & \text{where}~\lambda \vec x.~c = r.\action, \vec d = r.\data \\
      & \text{where}~r = \argmin_i\{\satisfy \sigma \pkt \mu b_i\} \\
      & \text{where}~b_i = \sigma(t.\id)[i].\matches = t.\keys \\
      \{(\mu,\pkt)\}, & \mathit{otherwise}
    \end{cases}
  \end{array}\]
  \caption{Semantics of Commands}
  \label{fig:com-sem}
\end{figure}

@@

\begin{figure}
  \[\begin{array}{r>{\triangleq}cl}
  \multicolumn{3}{l}{(\cdot)[\sigma] : \Cmd \to \Cmd} \\
  \multicolumn{3}{l}{(\cdot)[\sigma] : \Prog \to \GCL} \\
  (h.f := e)[\sigma] && h.f := e   \\
  (\assume b)[\sigma] && \assume b \\
  (\assert b)[\sigma] && \assert b \\
  (c_1;c_2)[\sigma] && c_1[\sigma];c_2[\sigma] \\
  (c_1 \choice c_2)[\sigma] && c_1[\sigma] \choice c_2[\sigma] \\
  (\{c\})[\sigma] && \{c[\sigma]\} \\
  (t.\apply())[\sigma] &&
  \begin{array}[t]{l}
    (\\
    \quad \assume {r_1.\matches = t.\keys};\\
    \quad t.\action[r_1.\action](r_1.\data)\\
    )~\choice\\
    \multicolumn{1}{c}{\vdots}\\
    \choice~(\\
    \quad \assume {r_n.\matches = t.\keys};\\
    \quad t.\action[r_n.\action](r_n.\data)\\
    )\\
    \textrm{where}~[r_1;\ldots;r_n] = \sigma(t.\id)
  \end{array}
  \end{array}
  \]
  \caption{Table Elimination}
  \label{fig:table-elim}
\end{figure}

\begin{theorem}[Table Elimination]
  For every $c \in \Cmd$, well-formed state $\sigma : \Value \to \Row^*$, model
  $\mu \in \Model$, and packet $\pkt \in \Pkt$, then for any state $\tau :
  \Value \to \Row^*$,
  \[\denote{c}^\sigma~\mu~\pkt = \denote{c[\sigma]}^\tau~\mu~\pkt\]
\end{theorem}
\begin{proof}
\todo[inline]{By induction on the structure of $c$.}
\end{proof}

@@

The next step is to define a source to source translation that eliminates tables
in the absence of concrete table state, which we will write $c[?] \in \Cmd$.
This is the first step towards the goal of constraining control plane behavior.
Want the instrumented program $c[?]$ to fail on exactly the same situations that
$c[\sigma]$ does, however $c[?]$ is constrained to have no table applications.
The only remaining mechanism is to use $\assume$ and $\assert$ statements to
inspect $\sigma$. We define $c[?]$ in Figure~\ref{fig:table-instrument}.

@@

\begin{figure}[htp]
  \[\begin{array}{r >{\triangleq}cl}
  \multicolumn{3}{l}{(\cdot)[?] : \Cmd \to \Cmd} \\
  (h.f := e)[?] && h.f := e \\
  (\assume b)[?] && \assume b \\
  (\assert b)[?] && \assert b\\
  (c_1;c_2)[?] && c_1[?];c_2[?]\\
  (c_1 \choice c_2)[?] && c_1[?] \choice c_2[?]\\
  (\{c\})[?] && \{c[?]\}\\
  (t.\apply())[?] &&
  \begin{array}[t]{l} \displaystyle
    \{ \\
    \quad \rho.\matches := \overrightarrow\havoc; \\
    \quad \rho.\action := \havoc;\\
    \quad \rho.\data := \overrightarrow\havoc; \\
    \quad \assume \rho.\action < |t.\actions|; \\
    \quad \assume \rho.\matches = t.\keys; \\
    \quad \displaystyle \bigchoice_{i=0}^{|t.\action| -1}
    \left(\begin{array}{l}
    \assume \rho.\action = i;\\
    t.\action[i](\rho.\data)
    \end{array}\right)\\
    \} \choice ( \\
    \quad \assume \forall \rho \in t. \rho.\matches \neq t.\keys; \\
    \quad \SKIP \\
    )
  \end{array}
  \end{array}
  \]
  \caption{Symbolic Table instrumentation. Typing information is omitted for
    brevity; each $\havoc$ must have a appropriate bit-width annotation. For
    simplicity, we assume each action in a table has the same number (and types) of
    parameters.}
  \label{fig:table-instrument}
\end{figure}

\begin{figure}[htp]
  \[\begin{array}{r >{\triangleq}cl}
  \multicolumn{3}{l}{\state\sigma_c : (\Value \to \Row^*) \to \BExpr} \\
  \state\sigma_c = \bigwedge_{ t \in \tables(c) } \bigvee_{r_i \in \sigma(t.\id)}
  \rho = r_i
  \end{array}
  \]
  \caption{Symbolic State Encoding}
  \label{fig:table-instrument}
\end{figure}

%% \begin{lemma}[Models Invariant over Actions]
%%   \label{lem:models-invariant-over-actions}
%%   For a closed action command $c$, $\mu,\mu'
%%   \in \Model$, $\sigma \in \Value \to
%%   \Row^*$, $\pkt \in \Pkt$, there exists some $\pkt'$
%%   \[\denote{c}^\sigma\,\mu\,\pkt = \{\mu,\pkt'\}\]
%%   \end{lemma}

%% \begin{proof}
%%   Proceed by induction over $c$.
%%   \begin{enumerate}[align=left]
%%   \item[$(c = h.f = e)$] Immediate.
%%   \item[$(c = c_1;c_2)$]
%%     By IH($c_1$) there exists some $\pkt''$ such that
%%     $\denote{c_1}^\sigma\,\mu\,\pkt = \{(\mu,\pkt'')\}$.
%%     By IH($c_2$) there exists some $\pkt'$ such that
%%     $\denote{c_2}^\sigma,\mu\,pkt'' = \{(\mu, \pkt')\}$.
%%     The result follows by definition of $\denote{-}$.
%%     \end{enumerate}
%% \end{proof}

\begin{samepage}
\begin{theorem}[Instrumentation Equivalence]
  \label{thm:instr-equiv}
  For every $c \in \Cmd$, well-formed state $\sigma : \Value \to \Row^*$, model
  $\mu \in \Model$, and packet $\pkt \in \Pkt$, \[\denote{c}^\sigma~\mu~\pkt = \denote{c[?];\assume \state\sigma_c}^\sigma~\mu~\pkt.\]
\end{theorem}
\end{samepage}

\begin{proof}
  \todo[inline]{Proceed by induction on the structure of $c$, leaving $\sigma$, $\mu$, $\pkt$ general}

  %% \begin{enumerate}[align=left]
  %% \item[$(c = h.f := e)$]
  %%   Let $\sigma$,$\mu$,$\pkt$ be given.
  %%   By definition $h.f:=e[?] = h.f:=e$. The result follows.
  %% \item[$(c = \assume b)$] Similar. $[?]$ is no-op.
  %% \item[$(c = \assert b)$] Similar. $[?]$ is no-op.
  %% \item[$(c = c_1;c_2)$]
  %%   Let $\sigma,\mu,\pkt$ be given.
  %%   By definition $(c_1;c_2)[?] = c_1[?]; c_2[?]$.
  %%   Compute as follows:
  %%   \[\begin{array}{>{\displaystyle}ll}
  %%   \denote{(c_1;c_2)[?]}^\sigma~\mu~\pkt \\
  %%   = \denote{c_1[?];c_2[?]}^\sigma~\mu\pkt
  %%   & \text{by definition of } [?]  \\
  %%   = \bigcup_{(\mu',\pkt') \in \denote{c_1[?]}^\sigma\,\mu\,\pkt}\denote{c_2[?]}^\sigma~\mu'~\pkt'
  %%   & \text{by definition of }\denote{-} \\
  %%   = \bigcup_{(\mu',\pkt') \in \denote{c_1}^\sigma\,\mu\,\pkt}\denote{c_2[?]}^\sigma~\mu'~\pkt'
  %%   & \text{by } \mathit{IH}(c_1, \mu, \pkt) \\
  %%   = \bigcup_{(\mu',\pkt') \in \denote{c_1}^\sigma\,\mu\,\pkt}\denote{c_2}^\sigma~\mu'~\pkt'
  %%   & \text{by } \mathit{IH}(c_2, \mu', \pkt') \\
  %%   = \denote{c_1;c_2}^\sigma~\mu~\pkt
  %%   & \text{by definition of } \denote{-}
  %%   \end{array}
  %%   \]
  %% \item[$(c = c_1 \choice c_2)$] Similar, by IHs.

  %% \item[$(c = t.\apply())$]
  %%   Let $\sigma$,$\mu$,$\pkt$ be given.
  %%   Let $r_1,\ldots,r_n = \sigma(t)$.
  %%   There are two cases, either the packet hits a row, or hits no row. Explore each case separately.
  %%   \begin{enumerate}
  %%   \item[\textsc{Hit}]
  %%     \textitP{}

  %%   \item[\textsc{Miss}] In this case, there is no $r_i$ such that
  %%     $\satisfy\sigma\pkt\mu r_i.\matches = t.\keys$, which means that the LHS
  %%     of the toplevel choice operator should evaluate to $\{\}$ \todo{compute
  %%       this}, and the RHS evaluates to $\{\mu,\pkt\}$\todo{compute this}, which
  %%     is precisely the result we get when executing
  %%     $\denote{t.apply()}^\sigma~\mu~\pkt$.

  %%   \end{enumerate}

  %% \end{enumerate}
\end{proof}

\subsection{Ghost State}

Following p4v, we want construct a ghostly header $\gamma$ from a program $p$
and use it to instrument the program to capture all dataplane behaviors.

\[
\begin{array}{l>{\triangleq}cl}
  \multicolumn{3}{l}{\ghostly :: \Cmd \to \Cmd} \\
  \multicolumn{3}{l}{\ghostly :: \Prog \to \GCL} \\
  \ghostly(h.f := e)
  && h.f := e \\
  \ghostly(\assert b)
  && \assert b\\
  \ghostly(\assume b)
  && \assume b \\
  \ghostly(c_1;c_2)
  && \ghostly(c_1);\ghostly(c_2) \\
  \ghostly(c_1\choice c_2)
  && \ghostly(c_1) \choice \ghostly(c_1)\\
  \ghostly(t.\apply())
  &&\\
  \multicolumn{3}{c}{
    \begin{array}[t]{>{\displaystyle\qquad}l}
       \gamma.t_{\mathit{reach}} := [1]_1; \\
       \gamma.t_{\mathit{read}(0)} := t.\keys[0]; \\
       \vdots \\
       \gamma.t_{\mathit{read}(|t.\keys|-1)} := t.\keys[|t.\keys|-1]; \\[1em]
       \left(\gamma.t_{\mathit{hit}} := [0]_1 \right)
       \choice \left(\bigchoice_{j = 1}^{|t.\action|-1}{
         \left(\begin{array}{l}
           \gamma.t_{\mathit{hit}} := [1]_1 ;\\
           \gamma.t_{\mathit{action}} := [1]_1;\\
           t.\action[j](\overrightarrow{\gamma.t_{\mathit{data}}})
         \end{array}\right)
       }\right)
     \end{array}}
\end{array}
\]

And now what we want to prove is

>>>>

% record fields
\newcommand{\matches}{\mathsf{matches}}
\newcommand{\action}{\mathsf{action}}
\newcommand{\actions}{\mathsf{actions}}
\newcommand{\keys}{\mathsf{keys}}
\newcommand{\data}{\mathsf{data}}
\newcommand{\params}{\mathsf{params}}
\newcommand{\id}{\mathsf{id}}
\newcommand{\reach}{\mathit{\mathit{reach}}}
\newcommand{\hit}{\mathit{hit}}
\newcommand{\reads}{\mathit{reads}}

% commands
\newcommand{\assert}{\mathop{\mathsf{assert}}}
\newcommand{\assume}{\mathop{\mathsf{assume}}}
\newcommand{\apply}{\mathsf{apply}}
\newcommand{\choiceop}{\rotatebox[origin=c]{90}{$\sqsubset\!\!\!\sqsupset$}}
\newcommand{\choice}{\mathbin{\choiceop}}
\DeclareMathOperator*{\bigchoice}{\scalerel*{\choiceop}{\sum}}
\newcommand{\havoc}[1]{\mathop{\textsf{havoc}}#1}
\newcommand{\SKIP}{\mathsf{skip}}

% math
\newcommand{\argmin}{\mathop{\textrm{argmin}}}
\newcommand{\WP}{\textit{wp}}
\newcommand{\satisfy}[3]{({#1,#3}) \models_{#2}}
\newcommand{\notsatisfy}[3]{({#1,#3})\not\models_{#2}}
\newcommand{\nondet}{\mathsf{nondet}}
\newcommand{\ghostly}{\mathsf{ghostly}}
\newcommand{\state}[1]{\llparenthesis#1\rrparenthesis}
\newcommand{\instr}{[?]}
\newcommand{\ginstr}{[\mathghost]}
\newcommand{\qe}{\textsc{Qe}}

@@

\begin{figure}[htp]
  \[\begin{array}{lclll}
    \multicolumn{3}{l}{\Value} \\
    v & ::= & [n]_w  & \textit{Bitvector Literal} & (a,g,p,\iota) \\
    \multicolumn{3}{l}{\SymbRow} \\
    \rho & ::= & \left\{
    \begin{array}{l}
      \matches = \List[x];\\
      \action = x;\\
      \data = \List[x]
    \end{array}\right\} & \textit{Symbolic Row} & (\iota)\\
    \multicolumn{3}{l}{\BVExpr} \\
    e & ::= & v & \textit{Value} & (a,g,p,\iota)\\
      & \mid & h.f & \textit{Header Access} & (a,g,p,\iota)\\
      & \mid & \rho.\matches[i] & \textit{Symbolic Match} & (\iota)\\
      & \mid & \rho.\action & \textit{Symbolic Action} & (\iota) \\
      & \mid & \rho.\data[i] & \textit{Symbolic Action Data} & (\iota) \\
      & \mid & x & \textit{Action Data Variable} & (a^*) \\
      & \mid & e \binop e & \textit{Arbitrary Binary Operation} & (a,g,p,\iota) \\
    & \mid & \unop e & \textit{Arbitrary Unary Operation} & (a,g,p,\iota) \\
    \multicolumn{3}{l}{\Table} \\
    t & ::= & \left\{\begin{array}{l}
    t.\id = v; \\
    t.\keys = \List[h.f]; \\
    t.\actions = \List[\lambda \vec x.\,a]\\
    \end{array}\right\} & \textit{Table} & (p)\\
    \multicolumn{3}{l}{\BExpr} \\
    b & ::= & \FALSE  & \textit{Absurdity}& (a,g,p,\iota)\\
      & \mid & b \Rightarrow b & \textit{Implication} & (a,g,p,\iota)\\
    & \mid & e = e & \textit{Bitvector Equality} & (a,g,p,\iota)\\
    & \mid & \rho \in t & \textit{Table membership} & (\iota)\\
    & \mid & \forall \rho.~b & \textit{Row Universality} & (\iota)\\
    \multicolumn{4}{l}{c \in \Cmd, a \in \Action, g \in \GCL, p \in \Prog, \iota \in \Instr } \\
    c & ::=  & h.f := e & \textit{Assignment} & (g,p,\iota) \\
      %% & \mid & \havoc \rho & \textit{Row Havoc} & (\iota) \\
      & \mid & \assert b & \textit{Assertion} & (g,p,\iota)\\
      & \mid & \assume b & \textit{Assumption} & (g,p,\iota) \\
      & \mid & t.\apply() & \textit{Table Application} & (p)\\
      & \mid & c;c & \textit{Sequential Composition} & (a,g,p,\iota) \\
      & \mid & c \choice c & \textit{Nondeterministic Choice} & (g,p,\iota)\\
      %% & \mid & \{c\} & \textit{Scope} & (g,p,\iota)\\
    \multicolumn{3}{l}{
      \begin{array}{ll}
        x \in \Var & h \in \Hdr\\
        n,w,i \in \mathbb{N} & f \in \Field
    \end{array}} & \textit{Sets }\\
  \end{array}
\]
\caption{Grammar of Commands $c \in \Cmd$. Legal action indicated by $a \in
  \Action$, legal instrumented command variants indicated by $\iota \in \Instr$,
  dataplane programs indicated by $p \in \Prog$, and table-free commands
  indicated by $g \in \GCL$.}
\label{fig:grammar}
\end{figure}

@@

@@

A state map $\sigma : \Value \to \Row^*$ is well-formed for a command $c$ iff the
following conditions hold:
\begin{enumerate}
\item[\textsc{Total}] For every $t \in \tables(c)$ $\pkt \in \pkt$ there exists a row $r \in \sigma(t.\id)$ such that $r.matches = \pkt(t.keys)$.
\item[\textsc{WellFormedRows}] For every $t\in \tables(c)$, every row in $\sigma(t.\id)$ is well-formed.
\item[\textsc{NoShadow}] For every $t \in \tables(c)$, with $r_1,\ldots,r_n =
  \sigma(t.\id)$, for every $1 \leq i \leq n$, there does not exist $1 \leq j <
  i$ such that $r_i.\matches = r_j.\matches$.
\end{enumerate}
\paragraph{Row Models}
We also need models $\mu \in \Model$ to indicate which symbolic rows corresponds
to which concrete rows, i.e. $\Model = \SymbRow \rightharpoonup \Row$. We write
$\mu\{\rho \mapsto r\}$ to extend $\mu$ by mapping $\rho$ to $r$, and write $\mu
\setminus \rho$, to remove $\rho$ from the map.

@@

\begin{figure}[htp]
  \[\begin{array}{l >{\triangleq}cl}
  \multicolumn{3}{l}{\denote{c}^\sigma : \Model \to \Pkt \to \mathcal{P}(\Model \times \Pkt)_\error }\\
    \denote{h.f := e}^\sigma\,\mu\,\pkt && \{(\mu, \pkt\{h \mapsto v\} \mid v \in \edenote{e}\,\mu\,\pkt\})\} \\
    \denote{\assume b}^\sigma\,\mu\,\pkt &&
    \begin{cases}
      \{(\mu,\pkt)\} & \satisfy \sigma \pkt \mu b \\
      \{\} & \textit{otherwise} \\
    \end{cases}\\
    \denote{\assert b}^\sigma\,\mu\,\pkt &&
    \begin{cases}
      \{(\mu,\pkt)\} & \satisfy \sigma \pkt \mu b \\
      \error & \textit{otherwise}
    \end{cases} \\
    %% \denote{\havoc \rho}^\sigma\,\mu\,\pkt &&
    %% \{(\mu\{\rho \mapsto r\}, \pkt) \mid r \in \Row\} \\
    \denote{c_1;c_2}^\sigma\,\mu\,\pkt&&
    \bigcup_{(\mu',\pkt') \in \denote{c_1}^\sigma\,\mu\,\pkt}\denote{c_2}^\sigma\,\mu'\,\pkt' \\
    \denote{c_1 \choice c_2}^\sigma\,\mu\,\pkt &&
    \denote{c_1}^\sigma\,\mu\,\pkt \cup \denote{c_2}^\sigma\,\mu\,\pkt \\
    %% \denote{\{c\}}^\sigma\,\mu\,\pkt &&
    %% \{(\mu,\pkt') \mid (\mu',\pkt') \in \denote{c}^\sigma\,\mu\,\pkt \}\\
    \denote{t.\apply()}^\sigma\,\mu\,\pkt &&
    \begin{cases}
      \denote{c[\vec d/\vec x]}^\sigma\,\mu\,\pkt,
      & \text{where}~\lambda \vec x.~c = r.\action, \vec d = r.\data \\
      & \text{where}~r = \argmin_i\{\satisfy \sigma \pkt \mu b_i\} \\
      & \text{where}~b_i = \sigma(t.\id)[i].\matches = t.\keys \\
      \{(\mu,\pkt)\}, & \mathit{otherwise}
    \end{cases}
  \end{array}\]
  \caption{Semantics of Commands}
  \label{fig:com-sem}
\end{figure}

@@

\begin{figure}
  \[\begin{array}{r>{\triangleq}cl}
  \multicolumn{3}{l}{(\cdot)[\sigma] : \Cmd \to \Cmd} \\
  (h.f := e)[\sigma] && h.f := e   \\
  (\assume b)[\sigma] && \assume b \\
  (\assert b)[\sigma] && \assert b \\
  (c_1;c_2)[\sigma] && c_1[\sigma];c_2[\sigma] \\
  (c_1 \choice c_2)[\sigma] && c_1[\sigma] \choice c_2[\sigma] \\
  %% (\{c\})[\sigma] && \{c[\sigma]\} \\
  (t.\apply())[\sigma] &&
  \begin{array}[t]{l}
    (\\
    \quad \assume {r_1.\matches = t.\keys};\\
    \quad t.\action[r_1.\action](r_1.\data)\\
    )~\choice\\
    \phantom)~\,\vdots\\
    \phantom)\,\choice~(\\
    \quad \assume {r_n.\matches = t.\keys};\\
    \quad t.\action[r_n.\action](r_n.\data)\\
    )\choice (\\
    \quad \displaystyle \assume {\bigwedge_{i=1}^n r_n.\matches \neq t.\keys};\\
    \quad \SKIP \\
    )\\
    \textrm{where}~[r_1;\ldots;r_n] = \sigma(t.\id)
  \end{array}
  \end{array}
  \]
  \caption{Table Substitution}
  \label{fig:table-elim}
\end{figure}

\begin{theorem}[Table Substitution]
  For every $c \in \Cmd$, well-formed state $\sigma : \Value \to \Row^*$, model
  $\mu \in \Model$, and packet $\pkt \in \Pkt$, then for any state $\tau :
  \Value \to \Row^*$,
  \[\denote{c}^\sigma~\mu~\pkt = \denote{c[\sigma]}^\tau~\mu~\pkt\]
\end{theorem}
\begin{proof}
\todo[inline]{By induction on the structure of $c$.}
\end{proof}

@@

The next step is to define a source to source translation that eliminates tables
in the absence of concrete table state, which we will write $c\instr \in \Cmd$.
This is the first step towards the goal of constraining control plane behavior.
Want the instrumented program $c\instr$ to fail in exactly the same situations that
$c\instr$ does, however $c\instr$ is constrained to have no table applications.
The only remaining mechanism is to use $\assume$ and $\assert$ statements to
inspect $\sigma$. We define $c[?]$ in Figure~\ref{fig:table-instrument}.

@@

We augment the headers for a program $c$ with ghost headers $\gamma_t$ for each
table $t \in \tables(c)$. The header $\gamma_t$ has defined fields
$\reach$, which is set to true if the packet reaches table $t$,
$\hit$ which is set to true the packet hits in the table, and
$\reads$ which is really a list of $|t.\keys|$ variables that stores the
values of the packet when evaluating $t.\keys$.

\begin{figure}[htp]
  %% \begin{minipage}[t]{0.5\textwidth}
  %% \[\begin{array}{r >{\triangleq}cl}
  %% \multicolumn{3}{l}{(\cdot)\instr : \Cmd \to \Cmd} \\
  %% (h.f := e)\instr && h.f := e \\
  %% (\assume b)\instr && b \\
  %% (\assert b)\instr && \assert b\\
  %% (c_1;c_2)\instr && c_1\instr;c_2\instr\\
  %% (c_1 \choice c_2)\instr && c_1\instr \choice c_2\instr\\
  %% (\{c\})\instr && \{c\instr\}\\
  %% (t.\apply())\instr && \\
  %% \multicolumn{3}{r}{
  %%   \qquad
  %%   \begin{array}[t]{l} \displaystyle
  %%     \{ \\
  %%     \quad \havoc \rho;\\
  %%     \quad \assume \rho \in t; \\
  %%     \quad \assume \rho.\matches = t.\keys; \\
  %%     \quad \displaystyle \bigchoice_{i=0}^{|t.\action| -1}
  %%     \left(\begin{array}{l}
  %%       \assume \rho.\action = i;\\
  %%       t.\action[i](\rho.\data)
  %%     \end{array}\right)\\
  %%     \} \choice ( \\
  %%     \quad \assume \forall \rho. \rho \in t \Rightarrow \rho.\matches \neq t.\keys;\\
  %%     \quad \SKIP \\
  %%     )
  %%   \end{array}}
  %% \end{array}
  %% \]
  %% \end{minipage}\begin{minipage}[t]{0.45\textwidth}
\[\begin{array}{r >{\triangleq}cl}
  \multicolumn{3}{l}{(\cdot)\instr : \Cmd \to \Cmd} \\
  (h.f := e)\instr && h.f := e \\
  (\assume b)\instr && \assume b \\
  (\assert b)\instr && \assert b\\
  (c_1;c_2)\instr && c_1\instr;c_2\instr\\
  (c_1 \choice c_2)\instr && c_1\instr \choice c_2\instr\\
  (t.\apply())\instr &&\\
  \multicolumn{3}{l}{
    \qquad
    \begin{array}[t]{l} \displaystyle
    \assume \rho_t.\matches = t.\keys; \\
    \displaystyle \bigchoice_{i=0}^{|t.\action| -1}
    \left(\begin{array}{l}
    \assume \rho_t.\action = i;\\
    t.\action[i](\rho_t.\data)
    \end{array}\right)\\
  \end{array}}
  \end{array}
  \]
  \caption{Symbolic Table instrumentation. For simplicity, we
    assume each action in a table has the same number (and types) of parameters}
  \label{fig:table-instrument}
\end{figure}

\begin{figure}[htp]
  \[\begin{array}{r cl}
  \multicolumn{3}{l}{\state\sigma_c : (\Value \to \Row^*) \to \BExpr} \\
  \state\sigma_c &\triangleq
  & \displaystyle\bigwedge_{ t \in \tables(c) }
  \bigvee_{r_i \in \sigma(t.\id)} \rho = r_i
  \end{array}
  \]
  \caption{Symbolic State Encoding}
  \label{fig:table-instrument}
\end{figure}

\begin{samepage}
\begin{theorem}[Instrumentation Equivalence]
  \label{thm:instr-equiv}
  For every $c \in \Cmd$, well-formed state $\sigma : \Value \to \Row^*$, model
  $\mu \in \Model$, and packet $\pkt \in \Pkt$, \[\denote{c}^\sigma~\mu~\pkt = \denote{\assume \state\sigma_c; c\instr}^\sigma~\mu~\pkt.\]
\end{theorem}
\begin{proof}
\todo[inline]{do the proof}
\end{proof}
\end{samepage}

\subsection{Predicate Semantics}

Now, we would like to be able to translate the result from
Theorem~\ref{thm:instr-equiv} into a
symbolic representation. To do this, we'll use the weakest precondition, defined
in Figure~\ref{fig:wp}

\begin{figure}[H]
  \[
  \begin{array}{r >{\triangleq}c l}
    \WP(h.f := e, b)
    && b[e/h.f] \\
    %% \WP(\havoc \rho, b)
    %% && \forall \rho.\,b \\
    \WP(\assert b_a, b)
    && b_a \wedge b \\
    \WP(\assume b_a, b)
    && b_a \Rightarrow b \\
    \WP(t.\apply,b) && \textrm{\textcolor{red}{undefined}} \\
    \WP(c_1;c_2, b)
    && \WP(c_1,\WP(c_2,b)) \\
    \WP(c_1 \choice c_2, b)
    && \WP(c_1,b) \wedge \WP(c_2,b)
  \end{array}
  \]
  \caption{Dijkstra's weakest precondition function}
  \label{fig:wp}
\end{figure}

Now, we need to relate states $\sigma$ and the computed $\WP(c\instr,\TRUE)$.
Note that since there are no havocs introduced in $c\instr$, all symbolic rows
$\rho$, and header values $h.f$ are unbound.

This means that we must check the validity of \[\state\sigma_c \Rightarrow
\WP(c\instr,\TRUE)\]

\clearpage
\section{The Research Problem}
